From Reversible Programs to \\ Univalent Universes and Back

### 2021-06-16 Pi+/paper/finite.tex
Remarks in sec 4.3
Some work on explaining immersion
Explain groups and deloopings

@@

\subsection{Permutations}~\label{subsec:permutations}

@@

@@

\begin{proposition}
  Let $\eta : A \to F(A)$ be given by $a \mapsto \inl(a) \cons \nil$. Given any group $G$ with a map $f : A \to G$,
  there is a canonical lifting of $f$ to a group homomorphism $\extend{f} : F(A) \to G$, such that
  $\extend{f} \comp \eta_{A} \htpy f$, that is, the type of group homomorphisms $h : \Grp(F(A),G)$ satisfying
  $h \comp \eta_{A} \htpy f$ is contractible. Equivalently, composition by $\eta_{A}$ is an equivalence
  $\Grp(F(A),G) \eqv (A \to G)$.
\end{proposition}

@@

\todo{Universal property}

@@

\begin{definition}[$\cox$]
  \begin{align*}
    \cancel
    & : \forall n \to (n \cons n \cons \nil) \cox \nil \\
    \swap
    & : \forall k, n \to (\suc[k] < n) \to (n \cons k \cons \nil) \cox (k \cons n \cons \nil) \\
    \braid
    & : \forall k, n \to (\suc[n] \cons k \cons \suc[n] \cons \nil) \cox (n \cons \suc[k] \cons n \cons \nil) \\
  \end{align*}
\end{definition}

@@

\todo{immersion}

\begin{theorem}
  For all $n : \Nat$,
  \[
    \Sn \eqv \Lehmer[n]
  \]
\end{theorem}

>>>>

\todo{Big example: Start from a listed permutation, show Lehmer code, then adjacent swaps.}
\todo{Justification for why we need group theory.}

@@

\subsection{Groups}

From universal algebra, a group is simply a set with a 0-ary constant, the neutral element, a binary operation for group
multiplication, and a unary inverse operation. A simple example is $\mathbb{Z}/n\mathbb{Z}$, where the neutral element
is 0, the inverse of $k$ is $-k$, and the group multiplication is given by addition modulo $n$. The neutral element has
to satisfy unit and inverse laws, and the multiplication has to be associative.

In type theory, a group $G$ can be defined as an $\hSet$ $S$ with the following pieces of data:

\begin{enumerate}
  \item a unit or neutral element $e : S$
  \item a multiplication function $m : S \times S \to S$ written as $(g_{1}, g_{2}) \mapsto g_{1} \mult g_{2}$, that satisfies
  \begin{enumerate}
    \item the unit laws, for all $g : S$, that \( g \mult e \id g \) and \( e \mult g \id g \)
    \item the associativity law, for all $g_{1}, g_{2}, g_{3} : S$, that \( g_{1} \mult (g_{2} \mult g_{3}) \id (g_{1} \mult g_{2}) \mult g_{3} \)
  \end{enumerate}
  \item an inversion function $i : S \to S$ written as $g \mapsto \inv{g}$, that satisfies
  \begin{enumerate}
    \item the inverse laws, for all $g : S$, that \( g \mult \inv{g} \id e \) and \( \inv{g} \mult g \id e \)
  \end{enumerate}
\end{enumerate}

However, more conveniently, in HoTT, we can instead use groupoids to talk about groups. A group can be identified with a
1-object groupoid, using a technique called delooping. The delooping of a group $G$ is a groupoid $\B{G}$ given by a
unique object $\pt$ with self-loops that are 1-paths $\pt \id_{\B{G}} \pt$ corresponding to the elements of $G$. Note
that the group operations are given by operations on the identity type automatically, with $\refl_{\pt}$ for the neutral
element, path composition for the group multiplication, and path inverse for the group inverse. These satisfy the group
laws as well, up to the identity type, using the groupoid coherence laws. Moreover, for 1-groups which are supposed to
be sets, these 2-paths should be propositions, so we have to restrict $\B{G}$ to be a 1-groupoid. Hence, a group is
simply given by a pointed, connected 1-type~\cite*{buchholtzHigherGroupsHomotopy2018,symmetryBook2021}.

For example, given a pointed type $(A:\UU, a:A)$, the automorphism group structure at $a$ is given by $a \id_{A} a$. Of
course, for 1-groups we will require that $a \id_{A} a$ is an $\hSet$, which is enforced by having $A$ be a groupoid. In
our running example for the permutation group on finite sets, we have that
$\Aut[\Fin[n]] \defeq (\Fin[n] \eqv \Fin[n]) \eqv (\Fin[n] \id_{\UU} \Fin[n])$, and the corresponding pointed, connected
groupoid is given by the classifying type $\UFin[n] \defeq \BAut[\Fin[n]]$, with
$\loopspace[\BAut[\Fin[n]],F_{n}] \eqv \Aut[\Fin[n]]$ as the corresponding automorphism group. \vc{revisit}

Alternatively, groups can also be presented using generators and relations, or a quotient of the free group \ldots
\subsection{Permutations}~\label{subsec:permutations}

@@

\todo{Free 1-group, needs to be truncated}

@@

\begin{proposition}
  Let $\eta_{A} : A \to F(A)$ be given by $a \mapsto \inl(a) \cons \nil$. Given any group $G$ with a map $f : A \to G$,
  there is a canonical lifting of $f$ to a group homomorphism $\extend{f} : F(A) \to G$, such that
  $\extend{f} \comp \eta_{A} \htpy f$, that is, the type of group homomorphisms $h : \Grp(F(A),G)$ satisfying
  $h \comp \eta_{A} \htpy f$ is contractible. Equivalently, composition by $\eta_{A}$ is an equivalence
  $\Grp(F(A),G) \eqv (A \to G)$.
\end{proposition}

\todo{diagram}
\todo{Plus/Minus}

@@

\todo{Universal property}
\todo{Examples: empty relation, van Kampen of $\pi_{1}$}

@@

\begin{definition}[$\cox$]
  \begin{align*}
    \cancel
    & : \forall n \to (n \cons n \cons \nil) \cox \nil \\
    \swap
    & : \forall k, n \to (\suc[k] < n) \to (n \cons k \cons \nil) \cox (k \cons n \cons \nil) \\
    \braid
    & : \forall n \to (\suc[n] \cons n \cons \suc[n] \cons \nil) \cox (n \cons \suc[n] \cons n \cons \nil) \\
  \end{align*}
\end{definition}

@@

\begin{definition}
  \[
    \encode{\List} : \List[\Fin[n]] \to \Lehmer[n]
  \]
\end{definition}

\begin{proposition}
  \leavevmode
  \begin{enumerate}
    \item \( l_{1} \cox* l_{2} \to \encode{\List}(l_{1}) \id \encode{\List}(l_{1}) \)
  \end{enumerate}
\end{proposition}

\begin{definition}
  \begin{align*}
    \encodeSn & : \Sn[n] \to \Lehmer[n] \\
    \decodeSn & : \Lehmer[n] \to \Sn[n]
  \end{align*}
\end{definition}

\begin{proposition}
  For all $n : \Nat$, $(\encodeSn, \decodeSn)$ is an equivalence.
\end{proposition}

\begin{corollary}
  For all $n : \Nat$,
  \(
    \Sn \eqv \Lehmer[n] \eqv \Aut[\Fin[n]]
  \).
\end{corollary}
